### 2020-08-20 analysis/paper/KWL-chronology-modeling.Rmd

---
title: "Supplementary Data: 14C Modeling for Kiwualn Chronology"
author: "Liying Wang and Ben Marwick"
date: "8/18/2020"
output: html_document
bibliography: references.bib
---

# Introduction

This document provides data and code for modeling the chronology of the upper component of Kiwulan. The full discussion of chronology is in the paper: The paper Standardization of ceramic shape: A case study from the Iron Age pottery from northeastern Taiwan. In this document, we calibrated all C14 dates [@Chen2007] first and then further modeled different phases related to foreign influences, pre- and post- European periods. Our purpose is to answer whether the C14 dates support a time range that corresponds to historical events of the presence of Europeans in Taiwan. The methods and results are included below.

```{r load-data}
kw_dates <- readxl::read_excel(here::here("analysis", "data", "raw_data", "KWL_C14_dates.xlsx"))
```

```{r tidy-data}
library(tidyverse)

# subset dates with error value
kw_dates_with_error <-
  kw_dates  %>%
  filter(!is.na(se))

# convert rce to numeric
kw_dates_with_error$rce <- as.numeric(kw_dates_with_error$rce)

# make some kind of ID
kw_dates_with_error$id <- with(kw_dates_with_error,
                               paste0(Square,"-",Layer))
```

# Calibration

We calibrated 23 dates from different contexts, including normal deposit, burials, and middens, at Kiwulan using oxcAAR package, which executes OxCal v4.4.1 with radiocarbon calibration curve IntCal20 [@Reimer2020].

```{r calibrate-C14data}
# Use oxcAAR to calibrate, run the following two lines to download ocxAAR 
# library(devtools)
# install_github('ISAAKiel/oxcAAR')
library(oxcAAR)
quickSetupOxcal()
date_kwl <- oxcalCalibrate(kw_dates_with_error$rce,
                           kw_dates_with_error$se,
                           kw_dates_with_error$`Lab code`)
date_kwl
plot(date_kwl)
calcurve_plot(date_kwl)
```

```{r data-for-modeling}
# upper layer with estimation of phases
kw_dates_with_error_three_phases <-
  kw_dates_with_error %>%
  filter(rce < 600) %>%
  mutate(phase = rce) %>%
  mutate(phase = ifelse(rce > 324, "Pre European", phase),
         phase = ifelse(rce < 324, "Post European", phase)) %>%
  select(`Lab code`, "rce", "se", "phase")
```

# Chronological model

We then constructed a Bayesian radiocarbon calibration model to obtain the range of pre-European and post-European we examined. We assume the phases are in an order that represents before and after the European arrival in 1624 [@Andrade2007]. The end of post-European is identified by the start of the Chinese migration to Yilan in 1780s [@Chen2007]. The results suggest that pre-European started between 1331–1635 cal AD at a 95.4% probability range and ended between 1458–1670 cal AD at a 99.7% probability range. The end date of pre-European also represents the start date of post-European, while the end date of post-European is 1517-1753 cal AD at a 95.4% probability range. The overall model is consistent with the historical data.

```{r modeling-upper-layer}
# contiguous phases, not overlapping
three_phases_con <- 'Plot()
{
    Sequence("Upper Layer")
    {
      Boundary("Start of Pre-European");
      Phase()
      {
        R_Date("NTU-3791", 340, 30);
        R_Date("NTU-4292", 510, 75);
        R_Date("NTU-4310", 360, 100);
        R_Date("NTU-4320", 340, 100);
        Interval( "Duration of Pre-European");
      };
        Boundary("Pre-European - Post-European");
      Phase()
      {
        R_Date("NTU-3993", 250, 40);
        R_Date("NTU-4016", 270, 45);
        R_Date("NTU-4311", 310, 100);
        R_Date("NTU-4419", 280, 70);
        Interval("Duration of Post-European");
      };
      Boundary("End of Post-European");
    };
};
'
```

```{r plot}
three_phases_con_results <- executeOxcalScript(three_phases_con)
three_phases_con_text <- readOxcalOutput(three_phases_con_results)
three_phases_result_con_data <- parseOxcalOutput(three_phases_con_text, only.R_Date = F)
str(three_phases_result_con_data)
print(three_phases_result_con_data)
plot(three_phases_result_con_data)
```