### 2020-08-22 analysis/paper/KWL-chronology-modeling.Rmd
updated chronology modeling file

---
title: "Supplementary Data: Radiocarbon Age Modeling for Kiwulan Chronology"
author: "Li-Ying Wang and Ben Marwick"
date: "8/18/2020"
output: html_document
bibliography: references.bib
---

# Introduction

This document provides data and code for calibrating radiocarbon dates and modeling the chronology of the upper component of Kiwulan. The full discussion of chronology is in the paper: Standardization of ceramic shape: A case study from the Iron Age pottery from northeastern Taiwan. There are 17 radiocarbon dates from the upper component according to the excavation report [@Chen2007]. We calibrated 11 radiocarbon dates after removing 6 dates that are not informative due to their long time ranges without standard deviation (< 200 yr). We then modeled two phases related to the European influence, pre- and post- European periods. Our purpose is to answer whether the radiocarbon dates support a time range that corresponds to our artifact-based chronology in terms of the presence of Europeans in Taiwan.

```{r load-data}
kw_dates <- readxl::read_excel(here::here("analysis", "data", "raw_data", "KWL_C14_dates.xlsx"))
```

```{r tidy-data}
library(tidyverse)

# subset dates with error value
kw_dates_with_error <-
  kw_dates  %>%
  filter(!is.na(se))

# convert rce to numeric
kw_dates_with_error$rce <- as.numeric(kw_dates_with_error$rce)

# make some kind of ID
kw_dates_with_error$id <- with(kw_dates_with_error,
                               paste0(Square,"-",Layer))

# upper layer with estimation of phases
kw_dates_with_error_three_phases <-
  kw_dates_with_error %>%
  filter(Component == "upper") %>%
  mutate(phase = rce) %>%
  mutate(phase = ifelse(rce > 324, "Pre European", phase),
         phase = ifelse(rce < 324, "Post European", phase)) %>%
  select(`Lab code`, "rce", "se", "phase", Context)
```

# Calibration

We calibrated 11 dates from different contexts, including artifact-bearing deposit, burials, and middens, at Kiwulan using oxcAAR package, which executes OxCal v4.4.1 with radiocarbon calibration curve IntCal20 [@Reimer2020].

```{r calibrate-C14data}
# Use oxcAAR to calibrate, run the following two lines to download ocxAAR 
# library(devtools)
# install_github('ISAAKiel/oxcAAR')
library(oxcAAR)
quickSetupOxcal()
date_kwl <- oxcalCalibrate(kw_dates_with_error$rce,
                           kw_dates_with_error$se,
                           kw_dates_with_error$`Lab code`)
date_kwl
plot(date_kwl)
calcurve_plot(date_kwl)
```

# Chronological model

We constructed a Bayesian radiocarbon calibration model for the 11 dates to obtain the range of the pre-European and the post-European periods we examined. We assume that those two phases are in a time sequence with the arrival of the Europeans in 1624 as a break [@Andrade2007]. The end of the post-European period is identified by the start of the Chinese migration to Yilan in 1780s [@Chen2007].

The results suggest that the pre-European period started between 998–1386 cal AD at a 95.4% probability range and ended between 1496–1656 cal AD at a 95.4% probability range. The start date of the post-European period is the same as the end date of the pre-European period (1496–1656 cal AD). The end date of the post-European period is 1524-1770 cal AD at a 93.4% probability range. In general, the overall model for radiocarbon age relative to the European contact is consistent with our artifact-based chronology.

```{r modeling-upper-layer}
# contiguous phases, not overlapping
three_phases_con <- 'Plot()
{
    Sequence("Upper Layer")
    {
      Boundary("Start of Pre-European");
      Phase()
      {
        R_Date("NTU-3791", 340, 30);
        R_Date("NTU-4292", 510, 75);
        R_Date("NTU-4310", 360, 100);
        R_Date("NTU-4320", 340, 100);
        R_Date("NTU-4304", 600, 75);
        R_Date("NTU-4423", 610, 90);
        R_Date("NTU-4319", 920, 105);
        Interval( "Duration of Pre-European");
      };
        Boundary("Pre-European - Post-European");
      Phase()
      {
        R_Date("NTU-3993", 250, 40);
        R_Date("NTU-4016", 270, 45);
        R_Date("NTU-4311", 310, 100);
        R_Date("NTU-4419", 280, 70);
        Interval("Duration of Post-European");
      };
      Boundary("End of Post-European");
    };
};
'
```

```{r plot}
three_phases_con_results <- executeOxcalScript(three_phases_con)
three_phases_con_text <- readOxcalOutput(three_phases_con_results)
three_phases_result_con_data <- parseOxcalOutput(three_phases_con_text, only.R_Date = F)
# str(three_phases_result_con_data)
# print(three_phases_result_con_data)
plot(three_phases_result_con_data)
```